---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rjmc

<!-- badges: start -->
<!-- badges: end -->

The goal of rjmc is to implement an example of RJMCMC for nested AR models.

## Installation

You can install the development version of rjmc from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("llrebecca21/rjmc")
```

To run the vignette it is also recommended to install, load, and update the following libraries:

```r
library(bayesplot) # for plotting trace plots
library(ggplot2) # for altering plots produced by bayesplot
library(latex2exp) # for including latex in plot titles
library(knitr) # for creating nice looking tables
library(tip) # for plotting posterior probabilities in ggplot
library(tidyr) # for data wrangling 
```

## Vignette

A vignette is available that walks through the example seen here in more detail as well as 3 other scenarios for different initial values of `k`. To run the vignette, either take the .Rmd file from the github repo for `rjmc` package in the vignettes folder, or install the `rjmc` package as above but with the `build_vignettes = TRUE` option stated.\




## Example

We will walk through how to use this package to implement Example 1.2 from Robert and Casella's "Efficient Construction of reversible jump Markov chain Monte Carlo proposal distributions" (2003) and Example 2.7 in Godsill's "On the Relationship Between Markov Chain Monte Carlo Methods for Model Uncertainty" (2001).

First need to create a time series of length `maxT` with a known set of true AR coefficients. We will use the coefficients given in the Godsill, 2001 paper.

```{r Create Time Series}
# set the seed for reproducibility
set.seed(100)
# Create a time series using arima.sim
# length of time series
maxT = 1000
# AR coefficients
ar_c = c(0.9402,
         -0.43,
         0.4167,
         -0.4969,
         0.4771,
         -0.5010,
         0.0509,
         -0.2357,
         0.4024,
         -0.1549)
x = arima.sim(model = list("ar" = ar_c), n = maxT)
```

True `k` is then equal to `10`. In the vignette, we will show a few examples with different starting values of `k` to see how well this rjmcmc sampler can perform.

Now we can call the library and run an example using the `rjmcmc_nested()` function.

```{r example}
library(rjmc)
library(bayesplot)
library(ggplot2)
# Set the number of iterations
iter = 3000
# Set the highest value k can possible take
kmax = 20
# Set starting value of k
k = 10
# Run the rjmcmc nested example on the time series
# Choose starting k = 10
# Create a list of data frames to store our chains
chains = 5
ex_1 = vector(mode = "list", length = chains)
for(i in 1:chains){
  ex_1[[i]]=rjmcmc_nested(iter = iter,k = k,sig2 = 1,x = x,kmax = kmax)
}

# extract the labels names
lab = colnames(ex_1[[1]])
```


```{r}
# Plot the trace plots of all the AR(k) parameters
bayesplot::mcmc_trace(x = ex_1, pars = lab[-c(21:22)]) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Due to the many trace plots, it is difficult to determine where our burn-in should be. So, let us plot the trace plots again but restricting to only the AR(1) through AR(10) parameters.
```{r}
bayesplot::mcmc_trace(x = ex_1, pars = lab[1:12]) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Seems the burn-in for this particular example should be around 200. Now we will plot the trace plots again with the burn-in and standardize the y-axis limits across all AR(k) plots.

# Plots with burn-in removed
```{r plots-with-burnin, message=FALSE, warning=FALSE}
burnin = 50

# Trace plots of sig2 and k
bayesplot::mcmc_trace(x = ex_1, pars = lab[21:22]) + xlim(burnin+1,3000)

# Plot the trace plots of all the AR(k) parameters
bayesplot::mcmc_trace(x = ex_1, pars = lab[-c(21,22)]) + xlim(burnin+1,iter) +
  ylim(-1,1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# Table of the posterior probability
# table(ex_1[-c(1:burnin),22])/(iter - burnin)
# Trace plots of only AR(1) through AR(10)
bayesplot::mcmc_trace(x = ex_1, pars = lab[1:12]) + xlim(burnin+1,iter)+
  ylim(-1,1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Looking at just the AR(1) through AR(10) trace plots (in base plot) we can overlay them with a horizontal line of the true AR coefficients. 

# Plots of Bayes Estimator Residual with burn-in removed

```{r}
par(mfrow = c(2,5), mar = c(2,2,1,1))
for(d in 1:10){
  plot(ex_1[[1]][burnin:iter,d], type = "l", ylim = c(-1,1), col = "lightblue")
  #plot true ar coefficient value
  abline(h=ar_c[d], col = "black")
  #plot colmeans ar coefficient value
  #abline(h=mean(ex_1[burnin:iter,d]), col = "blue")
}
```

```{r}
# Look at the values of the red and blue lines and store into a small matrix:
# true vs.colmeans
# create padded version of ar_c
true_ar = append(ar_c, rep(0,kmax-length(ar_c)))
comp_ar = matrix(NA, nrow = kmax, ncol = 4)
# fill first column with true_ar values
comp_ar[,1] = true_ar
for(i in 1:20){
  comp_ar[i,2] = mean(ex_1[[1]][burnin:iter,i])
  comp_ar[i,3] = sd(ex_1[[1]][burnin:iter,i])
  # The third column of the matrix contains the absolute value difference between the first two columns
  comp_ar[i,4] = (comp_ar[i,2] - comp_ar[i,1])/comp_ar[i,3]
}
comp_ar
plot(comp_ar[,4], ylab = "Standardized Residual",
     xlab = "k",
     xaxt = "n",
     las = 1,
     main = "Standardized Bayes Estimator Residual of\n the AR Coefficients")
abline(h = 0, col = "red")
axis(1, at = 1:20)


```
